---
title: "Taboo: Behavioral analysis"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(forcats)
library(lme4)
```


# Load data

```{r}
# read data
## setwd("/Users/seantrott/Dropbox/UCSD/Research/Ambiguity/Taboo/Experiment1/analysis/reports")
df = read.csv("../../data/processed/behavioral_results_public.csv")

```

## Remove sessions 10 and 15

```{r}
df = df %>%
  filter((session_id %in% c(10, 15)) == FALSE)

nrow(df)
length(unique(df$session_id))

df = df %>%
   filter(Native_Speaker == "Yes")

nrow(df)
```


## Descriptive statistics

```{r}
nrow(df)
table(df$session_id)
table(df$pair_type, df$trial_result)

table(df$Order)

length(unique(df$session_id))
length(unique(df$ppt_id))

## How many trials per session?
df_session = df %>%
  group_by(session_id, Order) %>%
  summarise(count = n())
mean(df_session$count)
median(df_session$count)
sd(df_session$count)
table(df_session$Order)
```

### Check data


```{r}
df_demo = df %>%
  group_by(ppt_id, Played_before, Gender, Native_Speaker) %>%
  summarise(vocab = mean(MINT_Score_Uncued_Correct),
            taboo = mean(Taboo_expertise))

table(df_demo$Native_Speaker)
table(df_demo$taboo)

df_demo %>%
  group_by(Native_Speaker) %>%
  summarise(mean_mint = mean(vocab),
            count = n())

```



### Participant-level characteristics

People scored very high on the MINT in general; the lowest score was 19.

```{r}
df_demo = df %>%
  group_by(ppt_id, Played_before, Gender) %>%
  summarise(vocab = mean(MINT_Score_Uncued_Correct),
            taboo = mean(Taboo_expertise))

table(df_demo$Gender)

df_demo %>%
  ggplot(aes(x = vocab)) +
  geom_histogram() +
  theme_minimal()

df_demo %>%
  ggplot(aes(x = taboo)) +
  geom_histogram() +
  theme_minimal()

table(df_demo$Played_before)
mean(df_demo$taboo)
sd(df_demo$taboo)
```


### Partner closeness scores

Strangers were automatically assigned a closeness score of 0.

```{r}
df %>%
  ggplot(aes(x = pair_type,
             y = partner_closeness)) +
  geom_violin() +
  theme_minimal()

df %>%
  group_by(pair_type) %>%
  summarise(closeness = mean(partner_closeness),
            sd_close = sd(partner_closeness))
```


# Modeling trial outcomes

There are three possible outcomes:

1) **Winning**: The listener successfully guesses the card.  
2) **Losing**: The speaker says a taboo word or gestures. (Or, for sessions 1-12, the speaker "skips" the card.)  
3) **Out of time**: The pair runs out of time on the card.

There are a few ways we might want to *model* these outcomes. We could group (2) and (3) together as a larger category of "Losing". Alternatively, we could analyze `Win` vs. `Lose` separately from `Win` vs. `Out of time`. The issue is complicated by the fact that speakers were allowed to skip cards for session 1-12; if $p(skip)$ differs as a function of `pair_type`, that might in turn affect the distribution of outcomes.

For each analysis, we will first conduct an analysis for **all sessions combined**, as well as the same analyses for only **sessions 12 and on**.

## Descriptive statistics

```{r dpi=300}
table(df$trial_result) / nrow(df)
table(df$pair_type, df$trial_result) / nrow(df)

df %>%
  ggplot(aes(x = pair_type,
             fill = trial_result)) +
  geom_bar(stat = "count", position = "dodge") +
  guides(fill=guide_legend(title="Trial result")) +
  labs(x = "Relationship type",
       y = "Count",
       title = "Trial outcomes by relationship type") +
  theme_minimal()


```


## Winning vs. out of time

### All sessions

First, we limit our analysis to trials that were either explicitly won or ran out of time.

```{r}
df_win_lose = df %>%
  filter(trial_result %in% c("Won", "Out of time"))

nrow(df_win_lose)
```

#### Note on random effects structure

Following Barr et al (2013), I begin with the **maximal** random effects structure including all random factors that could explain variance: `ppt_id`, `partner_id`, `Order`, `session_id`, and `deck` (`color`).

I then added a random slope for each random factor for the fixed effect of `pair_type`, which is the main variable we are interested in examining. Then, using nested model comparisons, I selected the random effects structure that was best justified by the data. 

In terms of random slopes, the only factor for which a random slope of `pair_type` explained significant variance was `Order` (i.e., whether participants played with their friend or with a stranger first; strangers were worse when that session began with strangers than when that session began with friends). All random intercepts tested explained additional variance.

```{r}
model_final_base = glmer(data = df_win_lose,
                              trial_result ~ 
                                (1 | ppt_id) +
                                (1 | partner_id) +
                                (1 + pair_type | Order) +
                                (1 | session_id) +
                                (1 | color),
                              control=glmerControl(optimizer="bobyqa"),
                              family = binomial())
```


#### Model comparisons

Now we add the covariates that might be important to account for.

It looks like the MINT score explains a very small amount of variance, and `trial_with_partner` explains significant amounts of variance. The probability of winning (relative to running out of time) increases with each trial one has with a partner.

```{r}
model_covariates = glmer(data = df_win_lose,
                          trial_result ~ 
                          Taboo_expertise +
                           MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_no_mint = glmer(data = df_win_lose,
                          trial_result ~ 
                          Taboo_expertise +
                           # MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_no_taboo = glmer(data = df_win_lose,
                          trial_result ~ 
                          # Taboo_expertise +
                           MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_no_turn_order = glmer(data = df_win_lose,
                          trial_result ~ 
                          Taboo_expertise +
                           MINT_Score_Uncued_Correct +
                           # trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())


anova(model_covariates, model_no_mint)
anova(model_covariates, model_no_taboo)
anova(model_covariates, model_no_turn_order)
summary(model_covariates)

```

#### Primary analyses

It appears that a fixed effect of `pair_type` does not explain significantly more variance than a model omitting this term. A model with an interaction between `pair_type` and `trial_with_partner` does explain marginally more variance (p = .18).

```{r}

model_interaction = glmer(data = df_win_lose,
                          trial_result ~ 
                            pair_type * trial_with_partner +
                           MINT_Score_Uncued_Correct +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        # control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_with_pair_type = glmer(data = df_win_lose,
                          trial_result ~ 
                            pair_type +
                           MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_final_base = glmer(data = df_win_lose,
                          trial_result ~ 
                           MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_no_turn = glmer(data = df_win_lose,
                          trial_result ~ 
                           MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

summary(model_interaction)
anova(model_interaction, model_with_pair_type)
anova(model_with_pair_type, model_final_base)
anova(model_with_pair_type, model_no_turn)
```


We can visualize this effect:

```{r}
df_win_lose = df_win_lose %>%
  mutate(
    outcome_numeric = case_when(
      trial_result == "Won" ~ 1,
      trial_result == "Out of time" ~ 0
    )
  )


df_win_lose %>%
  mutate(binned_trial = ntile(trial_with_partner, n = 10)) %>%
  group_by(binned_trial, pair_type) %>%
  summarise(success = mean(outcome_numeric),
            se = sd(outcome_numeric)/sqrt(n())) %>%
  ggplot(aes(x = binned_trial,
             y = success,
             color = pair_type)) +
  geom_smooth() +
  geom_errorbar(aes(ymin = success - 2*se, 
                  ymax = success + 2*se), 
              width=.2,
              position=position_dodge(.9)) +
  labs(x = "Binned Trial With Partner",
       y = "P(Win vs. Out of Time)",
       color = "Partner Type") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(breaks = c(1:10)) +
  theme_minimal() +
  theme(axis.title = element_text(size=rel(1.5)),
        axis.text = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.5)),
        strip.text.x = element_text(size = rel(1.5)),
        plot.title = element_text(size = rel(1.5)),
        legend.title = element_text(size = rel(1.5)))

ggsave("../../Figures/behavioral/win_vs_oot.png", dpi = 300)


```


## Winning vs. losing

First, we limit our analysis to trials that were either explicitly won or lost (produced a taboo word or gesture).

```{r}
df_win_lose2 = df %>%
  filter(trial_result %in% c("Won", "Lost"))

nrow(df_win_lose2)

df_win_lose2$trial_result = factor(df_win_lose2$trial_result, levels = c("Lost", "Won"))

table(df_win_lose2$trial_result)
```


#### Note on random effects structure

Following Barr et al (2013), I begin with the **maximal** random effects structure including all random factors that could explain variance: `ppt_id`, `partner_id`, `Order`, `session_id`, and `deck` (`color`).

I then added a random slope for each random factor for the fixed effect of `pair_type`, which is the main variable we are interested in examining. Then, using nested model comparisons, I selected the random effects structure that was best justified by the data. 

No random factors benefitted from a random slope for `pair_type`. Only random intercept for `session_id` explained significant variance, but for now, we live the other random intercepts in provided they don't intereferew ith  model fit.

```{r}

model_final_base = glmer(data = df_win_lose2,
                              trial_result ~ 
                                (1 | ppt_id) +
                                (1 | partner_id) +
                                (1 | Order) +
                                (1 | session_id) +
                                (1 | color),
                              control=glmerControl(optimizer="bobyqa"),
                              family = binomial())

```



#### Model comparisons

Now we add the covariates that might be important to account for.

Only `trial_with_partner` explains significant amounts of variance. Surprisingly, though, the probability of winning (relative to losing by using a taboo word or gesture) **decreases** with each trial one has with a partner.

```{r}
model_covariates = glmer(data = df_win_lose2,
                          trial_result ~ 
                          Taboo_expertise +
                           MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_no_mint = glmer(data = df_win_lose2,
                          trial_result ~ 
                          Taboo_expertise +
                           # MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_no_taboo = glmer(data = df_win_lose2,
                          trial_result ~ 
                          # Taboo_expertise +
                           MINT_Score_Uncued_Correct +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_no_turn_order = glmer(data = df_win_lose2,
                          trial_result ~ 
                          Taboo_expertise +
                           MINT_Score_Uncued_Correct +
                           # trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())


anova(model_covariates, model_no_mint)
anova(model_covariates, model_no_taboo)
anova(model_covariates, model_no_turn_order)
summary(model_covariates)

```


We can visualize this effect:

```{r}
df_win_lose2 = df_win_lose2 %>%
  mutate(
    outcome_numeric = case_when(
      trial_result == "Won" ~ 1,
      trial_result == "Lost" ~ 0
    )
  )


df_win_lose2 %>%
  mutate(binned_trial = ntile(trial_with_partner, n = 10)) %>%
  group_by(binned_trial, pair_type) %>%
  summarise(success = mean(outcome_numeric),
            se = sd(outcome_numeric)/sqrt(n())) %>%
  ggplot(aes(x = binned_trial,
             y = success,
             color = pair_type)) +
  geom_smooth() +
  geom_errorbar(aes(ymin = success - 2*se, 
                  ymax = success + 2*se), 
              width=.2) +
  labs(x = "Binned Trial With Partner",
       y = "P(Win vs. Lose)",
       color = "Partner Type") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(breaks = c(1:10)) +
  theme_minimal() +
  theme(axis.title = element_text(size=rel(1.5)),
        axis.text = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.5)),
        strip.text.x = element_text(size = rel(1.5)),
        plot.title = element_text(size = rel(1.5)),
        legend.title = element_text(size = rel(1.5)))

ggsave("../../Figures/behavioral/losing.png", dpi = 300)


```


#### Primary analyses

It appears that a fixed effect of `pair_type` does not explain significantly more variance than a model omitting this term, nor does an interaction between `pair_type * trial_with_partner`. 

```{r}

model_interaction = glmer(data = df_win_lose2,
                          trial_result ~ 
                            pair_type * trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 | Order) +
                          (1 | session_id) +
                          (1 | color),
                        # control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_with_pair_type = glmer(data = df_win_lose2,
                          trial_result ~ 
                            pair_type +
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

model_final_base = glmer(data = df_win_lose2,
                          trial_result ~ 
                           trial_with_partner +
                          (1 | ppt_id) +
                          (1 | partner_id) +
                          (1 + pair_type | Order) +
                          (1 | session_id) +
                          (1 | color),
                        control=glmerControl(optimizer="bobyqa"),
                        family = binomial())

summary(model_interaction)
anova(model_interaction, model_with_pair_type)
anova(model_with_pair_type, model_final_base)
```



