---
title: "Taboo: Linguistic analysis"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(forcats)
library(lme4)
library(corrplot)
library(broom.mixed)
```


# Load data

## Taboo data

```{r}
# setwd("/Users/seantrott/Dropbox/UCSD/Research/Ambiguity/Taboo/Experiment1/analysis/reports")
df = read.csv("../../data/processed/aligned_results_public.csv")
nrow(df)

table(df$session)
length(unique(df$session))

```


## Lexical statistics

```{r}
## Concreteness norms
df_concreteness = read_csv("../../data/lexical_statistics/brysbaert_norms.csv")
df_concreteness$word = df_concreteness$Word
## SUBTLEX frequency data
df_subtlex = read_csv("../../data/lexical_statistics/SUBTLEXusfrequencyabove1.csv")
df_subtlex$word = df_subtlex$Word
## AoA norms
df_aoa = read_csv("../../data/lexical_statistics/AoA_ratings_Kuperman_et_al_BRM.csv")
df_aoa$AoA = df_aoa$Rating.Mean
df_aoa$word = tolower(df_aoa$Word)
```


## Remove non-native speakers

```{r}
df = df %>%
  filter(Native_Speaker == "Yes")
nrow(df)
length(unique(df$ppt_id))
```


## Merge Taboo data with lexical statistics

### Merge with target concept

We then merge that data with the Taboo data to get `AoA`, `Frequency`, and `Concreteness` information about the target concept.

```{r}
df$word = tolower(df$card)
nrow(df)

df_merged = df %>%
  inner_join(df_concreteness, by = "word") %>%
  inner_join(df_subtlex, by = "word") %>%
  inner_join(df_aoa, by = "word") 
nrow(df_merged)


### Now remove critical NA observations for these columns
df_merged = df_merged %>% 
  drop_na(Conc.M) %>%
  drop_na(Lg10WF) %>%
  drop_na(AoA)
nrow(df_merged)


### Also restrict POS to Noun, Verb, and Adjective 
df_merged = df_merged %>%
  filter(Dom_Pos %in% c("Noun", "Verb", "Adjective"))
nrow(df_merged)
table(df_merged$Dom_Pos)
  
```


# Descriptive statistics

## Correlations between lexical statistics

```{r}
df_select_variables = df_merged %>%
  dplyr::select(Conc.M, Lg10WF, AoA)

cors = cor(df_select_variables, method='pearson')

corrplot(cors, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)


```


## Noun Phrases

Here, we create a subset of the data for expressions that *can conceivably* vary in length. 

```{r}
table(df_merged$RE_type)

df_just_nps = df_merged %>%
  filter(RE_type =="full NP")

df_just_nps %>%
  summarise(mean_length = mean(length),
            median_length = median(length),
            sd_length = sd(length))

nrow(df_just_nps)

```

What are the most common determiners / modifiers used for noun phrases that could vary in length?

```{r}
df_determiners = df_just_nps %>%
  group_by(det) %>%
  summarise(count = n(),
            prop = n() / nrow(df_just_nps))

head(df_determiners[order(-df_determiners$prop), ], 10)

df_determiners$rank = rank(-df_determiners$prop)

df_determiners %>%
  ggplot(aes(x = rank,
             y = log(count))) +
  geom_point() +
  labs(title = "Determiners: log(frequency) ~ Rank(frequency)") +
  theme_minimal()
```

What are the most common head nouns?

```{r}
df_heads = df_just_nps %>%
  group_by(head_noun) %>%
  summarise(count = n(),
            prop = n() / nrow(df_just_nps))

df_heads$rank = rank(-df_heads$prop)

head(df_heads[order(-df_heads$prop), ], 10)

df_heads %>%
  ggplot(aes(x = rank,
             y = log(count))) +
  geom_point() +
  labs(title = "Head nouns: log(frequency) ~ Rank(frequency)") +
  theme_minimal()
```


# Primary analysis 1: Does common ground influence the length of referring expressions?

First, we build a series of nested models. We consider only noun phrases that could conceivably vary in length (i.e., we exclude pronouns).

```{r}
model_full = glmer(data = df_just_nps,
                   length ~ 
                     pair_type +
                     trial_with_partner +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = poisson())


model_no_pair_type = glmer(data = df_just_nps,
                   length ~ 
                     # pair_type +
                     trial_with_partner +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = poisson())

model_no_twp = glmer(data = df_just_nps,
                   length ~ 
                     pair_type +
                     # trial_with_partner +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = poisson())

anova(model_full, model_no_pair_type)
anova(model_full, model_no_twp)

```


We can also visualize the coefficients from this full model:

```{r}
summary(model_full)
df_tidy = broom.mixed::tidy(model_full)

df_tidy %>%
  filter(effect == "fixed") %>%
  ggplot(aes(x = term,
             y = estimate)) +
  geom_point() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_errorbar(aes(ymin = estimate - 2*std.error, 
                    ymax = estimate + 2*std.error), 
                width=.2,
                position=position_dodge(.9)) +
  labs(x = "Predictor",
       y = "Estimate") +
  theme_minimal()
```


## Visualizations

```{r}

df_just_nps %>%
  ggplot(aes(x = pair_type,
             y = length,
             color = RE_type_granular)) +
  # geom_point(stat = "summary", fun.y = "mean") +
  stat_summary (fun = function(x){mean(x)},
              fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
              fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
              geom= 'pointrange', 
              position=position_dodge(width=0.95)) +
  labs(x = "Relationship",
       y = "Number of words",
       color = "RE Form") +
  theme_minimal()


df_just_nps %>%
  ggplot(aes(x = length)) +
  geom_histogram(bins = 8) +
  facet_wrap(~pair_type + RE_type_granular) +
  labs(x = "Number of words") +
  theme_minimal()
  
```



# Analysis 2: Does common ground influence referential form?

```{r}
df_restricted_form = df_merged %>%
  filter(RE_type %in% c('full NP', 'PRP 3rd'))
df_restricted_form$RE_type = factor(df_restricted_form$RE_type,
                                    levels = c("full NP", "PRP 3rd"))
nrow(df_restricted_form)

model_full = glmer(data = df_restricted_form,
                   RE_type ~ 
                     pair_type +
                     trial_with_partner +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_no_pair_type = glmer(data = df_restricted_form,
                   RE_type ~ 
                     # pair_type +
                     trial_with_partner +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_no_twp = glmer(data = df_restricted_form,
                   RE_type ~ 
                     pair_type +
                     # trial_with_partner +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

anova(model_full, model_no_pair_type)
anova(model_full, model_no_twp)

```


We can also visualize the coefficients from this full model:

```{r}
summary(model_full)

df_tidy = broom.mixed::tidy(model_full)

df_tidy %>%
  filter(effect == "fixed") %>%
  ggplot(aes(x = term,
             y = estimate)) +
  geom_point() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_errorbar(aes(ymin = estimate - 2*std.error, 
                    ymax = estimate + 2*std.error), 
                width=.2,
                position=position_dodge(.9)) +
  labs(x = "Predictor",
       y = "Estimate") +
  theme_minimal()
```


## Visualizations

```{r}
df_restricted_form %>%
  ggplot(aes(x = pair_type,
             fill = RE_type)) +
  geom_bar(stat = "count", position = "dodge") +
  theme_minimal()

df_restricted_form %>%
  # Set full NP = 0, PRP 3rd = 1
  mutate(re_numeric = as.numeric(RE_type) - 1) %>%
  ggplot(aes(x = pair_type,
             y = re_numeric)) +
  stat_summary (fun = function(x){mean(x)},
              fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
              fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
              geom= 'pointrange', 
              position=position_dodge(width=0.95)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Pair Type",
       y = "P(3rd-person pronoun)") +
  theme_minimal()


df_restricted_form %>%
  # Set full NP = 0, PRP 3rd = 1
  mutate(re_numeric = as.numeric(RE_type) - 1) %>%
  ggplot(aes(x = trial_with_partner,
             y = re_numeric,
             color = pair_type)) +
  stat_summary (fun = function(x){mean(x)},
              fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
              fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
              geom= 'pointrange', 
              position=position_dodge(width=0.95),
              alpha = .6) +
  labs(x = "Trial with partner",
       y = "P(3rd-person pronoun)") +
  guides(fill=guide_legend(title="Pair Type")) +
  scale_y_continuous(limits = c(0, 1)) +
  theme_minimal()
```





# Analysis 3: Do features of the RE interact with common ground to predict communicative success?

Here, we again limit our analysis to `full NPs` and `3rd-person pronouns`. 

```{r}
df_restricted_form = df_merged %>%
  filter(RE_type %in% c('full NP', 'PRP 3rd'))
df_restricted_form$RE_type = factor(df_restricted_form$RE_type,
                                    levels = c("full NP", "PRP 3rd"))
nrow(df_restricted_form)
table(df_restricted_form$RE_type)
```


## Win vs. Out of Time

First, we consider the contrast between `Win/Out of time`.

```{r}

df_win_oot = df_restricted_form %>%
  filter(trial_result != "Lost")
df_win_oot$trial_result = factor(df_win_oot$trial_result, levels = c("Out of time", "Won"))
nrow(df_win_oot)
table(df_win_oot$trial_result)

model_full = glmer(data = df_win_oot,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner * RE_type +
                     pair_type * length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_r1 = glmer(data = df_win_oot,
                   trial_result ~ 
                     pair_type + RE_type +
                     trial_with_partner * RE_type +
                     pair_type * length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_r2 = glmer(data = df_win_oot,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner + RE_type +
                     pair_type * length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_r3 = glmer(data = df_win_oot,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner * RE_type +
                     pair_type + length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_r4 = glmer(data = df_win_oot,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner * RE_type +
                     pair_type * length +
                     trial_with_partner + length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))



anova(model_full, model_r1)
anova(model_full, model_r2)
anova(model_full, model_r3)
anova(model_full, model_r4)

```


We can also visualize the coefficients from this full model:

```{r}
summary(model_full)
df_tidy = broom.mixed::tidy(model_full)

df_tidy %>%
  filter(effect == "fixed") %>%
  ggplot(aes(x = term,
             y = estimate)) +
  geom_point() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_errorbar(aes(ymin = estimate - 2*std.error, 
                    ymax = estimate + 2*std.error), 
                width=.2,
                position=position_dodge(.9)) +
  labs(x = "Predictor",
       y = "Estimate") +
  theme_minimal()
```


## Visualizations

```{r}
df_win_oot = df_win_oot %>%
  mutate(
    outcome_numeric = case_when(
      trial_result == "Won" ~ 1,
      trial_result == "Out of time" ~ 0
    )
  )

df_win_oot %>%
  ggplot(aes(x = pair_type,
             y = length,
             color = trial_result)) +
  stat_summary (fun = function(x){mean(x)},
              fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
              fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
              geom= 'pointrange', 
              position=position_dodge(width=0.95),
              alpha = .6) +
  labs(x = "Pair Type",
       y = "NP Length") +
  guides(color=guide_legend(title="Trial Result")) +
  theme_minimal()

df_win_oot %>%
  ggplot(aes(x = pair_type,
             y = outcome_numeric,
             color = RE_type)) +
  stat_summary (fun = function(x){mean(x)},
              fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
              fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
              geom= 'pointrange', 
              position=position_dodge(width=0.95),
              alpha = .6) +
  labs(x = "Pair Type",
       y = "P(success)") +
  guides(color=guide_legend(title="RE Type")) +
  theme_minimal()



```




## Win vs. Out of Time

First, we consider the contrast between `Win/Out of time`.

```{r}

df_win_lose = df_restricted_form %>%
  filter(trial_result != "Out of time")
df_win_lose$trial_result = factor(df_win_lose$trial_result, levels = c("Lost", "Won"))
nrow(df_win_lose)
table(df_win_lose$trial_result)

model_full = glmer(data = df_win_lose,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner * RE_type +
                     pair_type * length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))


model_r1 = glmer(data = df_win_lose,
                   trial_result ~ 
                     pair_type + RE_type +
                     trial_with_partner * RE_type +
                     pair_type * length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_r2 = glmer(data = df_win_lose,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner + RE_type +
                     pair_type * length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_r3 = glmer(data = df_win_lose,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner * RE_type +
                     pair_type + length +
                     trial_with_partner * length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))

model_r4 = glmer(data = df_win_lose,
                   trial_result ~ 
                     pair_type * RE_type +
                     trial_with_partner * RE_type +
                     pair_type * length +
                     trial_with_partner + length +
                     MINT_Score_Uncued_Correct +
                     Taboo_expertise +
                     Lg10WF +
                     AoA +
                     Conc.M +
                     Dom_Pos +
                    (1 | ppt_id) +
                    (1 | partner_id) +
                    (1 | Order) +
                    (1 | session_id) +
                    (1 | color),
                   family = binomial(),
                   control=glmerControl(optimizer="bobyqa"))



anova(model_full, model_r1)
anova(model_full, model_r2)
anova(model_full, model_r3)
anova(model_full, model_r4)
```


We can also visualize the coefficients from this full model:

```{r}
summary(model_full)
df_tidy = broom.mixed::tidy(model_full)

df_tidy %>%
  filter(effect == "fixed") %>%
  ggplot(aes(x = term,
             y = estimate)) +
  geom_point() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_errorbar(aes(ymin = estimate - 2*std.error, 
                    ymax = estimate + 2*std.error), 
                width=.2,
                position=position_dodge(.9)) +
  labs(x = "Predictor",
       y = "Estimate") +
  theme_minimal()
```


## Visualizations

```{r}
df_win_lose = df_win_lose %>%
  mutate(
    outcome_numeric = case_when(
      trial_result == "Won" ~ 1,
      trial_result == "Lost" ~ 0
    )
  )

df_win_lose %>%
  ggplot(aes(x = pair_type,
             y = length,
             color = trial_result)) +
  stat_summary (fun = function(x){mean(x)},
              fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
              fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
              geom= 'pointrange', 
              position=position_dodge(width=0.95),
              alpha = .6) +
  labs(x = "Pair Type",
       y = "NP Length") +
  guides(color=guide_legend(title="Trial Result")) +
  theme_minimal()

df_win_lose %>%
  ggplot(aes(x = pair_type,
             y = outcome_numeric,
             color = RE_type)) +
  stat_summary (fun = function(x){mean(x)},
              fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
              fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
              geom= 'pointrange', 
              position=position_dodge(width=0.95),
              alpha = .6) +
  labs(x = "Pair Type",
       y = "P(success)") +
  guides(color=guide_legend(title="RE Type")) +
  theme_minimal()



```



